The effect of prenatal environment on brain metabolic function and perception in a lizard

### 2025-07-30 ms.qmd
Update ms.qmd

#### Does prenatal CORT and temperature affect prey detection?
Models explained between 26 to 29.3 % of the variation in prey detection (see Table S1).

@@

## Discussion
We examined the impact of prenatal temperature and CORT on visual and chemical perception and determined whether metabolic function was the underlying reason for any perceptual differences we observed as has been shown in mammals [@hara_presynaptic_2014; @chaudhari2022early]. Our findings suggest that metabolic processes remained largely resilient to early environmental conditions. However, early-life conditions have region- and context-dependent effects on oxidative damage and prey detection. We observed interactive effects of prenatal CORT and incubation temperature on oxidative damage in the Optic Tecta, but no effects on perception of visual stimuli. In contrast, CORT decrease the time to perceived chemical stimuli despite the resilience of the Olfactory bulbs to prenatal conditions. Below, we discuss these findings in relation to mitochondrial metabolic function, oxidative stress, and behavioural performance.

>>>>

#### Does prenatal CORT and incubation temperature affect prey detection?
Models explained between 26 to 29.3 % of the variation in prey detection (see Table S1).

@@

## Discussion
We examined the impact of prenatal temperature and CORT on visual and chemical perception and determined whether metabolic function was the underlying reason for any perceptual differences we observed as has been shown in mammals [@hara_presynaptic_2014; @chaudhari2022early]. Our findings suggest that metabolic processes remained largely resilient to early environmental conditions. However, early-life conditions have region- and context-dependent effects on oxidative damage and prey detection. We observed interactive effects of prenatal CORT and incubation temperature on oxidative damage in the Optic Tecta, but no effects on perception of visual stimuli. In contrast, CORT decreased the time to perceive chemical stimuli despite the resilience of the Olfactory bulbs to prenatal conditions. Below, we discuss these findings in relation to mitochondrial metabolic function, oxidative stress, and behavioural performance.
